---
title: "TWFE models"
output: html_document
---

**Overview**

This file takes the dataset containing the individual birth records and the policy information and runs the traditional
TWFE analysis. It does it in a few different ways using individual vs aggregated to the state-month data.

**Inputs:** 

Note that the restricted births data is not available publicly and so is not uploaded to GitHub.
./restricted-data/births_all: this is the dataset of individual birth records merged to policy data

**Outputs:**

./home/corinneriddell/ACA-PTB/output-data/results.csv:Contains the parameter estimate and CI for the TWFE specifications

```{r set-path}
# set path
riddellPath <- c( "/data/riddell//R/x86_64-pc-linux-gnu-library/4.1/" , .libPaths() )
.libPaths(riddellPath)
```

```{r load-libraries}
# load libraries
library(magrittr)
library(tidyverse)
library(tibble)
library(readxl)
library(readr)
library(data.table)
library(ggrepel)
library(sandwich)
library(did)
library(tigris)

packageVersion("did")
```

```{r}
#load in the restricted births data
#all_years2 <- fread("./restricted-data/births_all")
all_years2 <- fread("/home/corinneriddell/ACA-PTB/restricted-data/births_all")
```

### DID model with individual level data

```{r}
# bare bones DID model
# individual level data, linear probability model
system.time(mod2 <- lm(CR_ptb ~ post_policy + state_r + as.factor(month_index), data = all_years2))
summary(mod2)

#robust standard errors
system.time(m2_var <- vcovHC(mod2, type="HC3")) #ten mins -- takes a long time when using the individual level data

lb = summary(mod2)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m2_var["post_policy","post_policy"]) 
ub = summary(mod2)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m2_var["post_policy","post_policy"])
c(lb, ub)

results <- data.frame(estimate = summary(mod2)$coefficients["post_policy", "Estimate"], 
           lb = lb, 
           ub = ub, 
           method = "individual-level TWFE LPM")
```

### DID model with aggregated state-month data

```{r}
all_years_agg <- all_years2 %>% 
  group_by(state_r, month_index) %>% 
  summarize(ptb_prop = mean(CR_ptb), 
            DOB_MM = first(DOB_MM), 
            DOB_YY = first(DOB_YY), 
            post_policy = first(post_policy), 
            ever_treated = first(ever_treated), 
            Expansion_Date = first(Expansion_Date)) %>%
  arrange(state_r, month_index)

all_years_agg %<>%
  ungroup() %>%
  group_by(state_r) %>%
  arrange(state_r, month_index) %>%
  mutate(time_since_policy = cumsum(post_policy))
  
mod5 <- glm(ptb_prop ~ post_policy + state_r + as.factor(month_index), data = all_years_agg, family = "gaussian") 
#Gaussian to keep things additive
summary(mod5)

#how does this estimate compare to model 2?

#robust SE
system.time(m5_var <- vcovHC(mod5, type="HC3"))

lb2 = summary(mod5)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m5_var["post_policy","post_policy"]) 
ub2 = summary(mod5)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m5_var["post_policy","post_policy"])
c(lb2, ub2)

results2 <- data.frame(estimate = summary(mod5)$coefficients["post_policy", "Estimate"], 
           lb = lb2, 
           ub = ub2, 
           method = "state-level TWFE LPM")

results2 <- bind_rows(results, results2)

#Below returns the same result of the previous specification so i commented it out
#this is an alternative specification but is equivalent to what is above
# mod5.2 <- lm(ptb_prop ~ post_policy + state_r + as.factor(month_index), data = all_years_agg) 
# summary(mod5.2)

# #robust SE
# system.time(m5_var2 <- vcovHC(mod5.2, type="HC3"))
# 
# lb3 = summary(mod5.2)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m5_var2["post_policy","post_policy"]) 
# ub3 = summary(mod5.2)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m5_var2["post_policy","post_policy"])
# 
# #double checking these are the same - they are 
# c(lb2, ub2)
# c(lb3, ub3)
```

### Run the alternative methods and see how these affect the results

```{r}
# TWFE if you only include those who eventually get the intervention 
mod6 <- glm(ptb_prop ~ post_policy + state_r + as.factor(month_index), 
            data = all_years_agg %>% filter(ever_treated == TRUE), family = "gaussian") #Gaussian to keep things additive
summary(mod6)

system.time(m6_var <- vcovHC(mod6, type="HC3"))

lb4 = summary(mod6)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m6_var["post_policy","post_policy"]) 
ub4 = summary(mod6)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m6_var["post_policy","post_policy"])
c(lb4, ub4)

results3 <- data.frame(estimate = summary(mod6)$coefficients["post_policy", "Estimate"], 
           lb = lb4, 
           ub = ub4, 
           method = "state-level TWFE ever-treated")

results3 <- bind_rows(results2, results3)
```

```{r}
#take away the states that make things unbalanced
all_years_agg2 <- all_years_agg %>% filter(! (state_r %in% c("AK", "MA", "MN", "MS", "VA")))

mod5_balanced <- glm(ptb_prop ~ post_policy + state_r + as.factor(month_index), data = all_years_agg2, family = "gaussian") 
#Gaussian to keep things additive
summary(mod5_balanced)

#how does this estimate compare to model 2?

#robust SE
system.time(m5_var <- vcovHC(mod5_balanced, type="HC3"))

lb2 = summary(mod5_balanced)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m5_var["post_policy","post_policy"]) 
ub2 = summary(mod5_balanced)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m5_var["post_policy","post_policy"])
c(lb2, ub2)


results4 <- data.frame(estimate = summary(mod5_balanced)$coefficients["post_policy", "Estimate"], 
           lb = lb2, 
           ub = ub2, 
           method = "state-level TWFE LPM balanced")

results4 <- bind_rows(results3, results4)

#add variables that are formatted as percentages (eg 0.02 becomes 2%)
results4 <- results4 %>% mutate(
  estimate2 = estimate*100, 
  lb2 = lb*100,
  ub2 = ub*100,
  description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")")
)

ggplot(results4, aes(y = method, x = estimate2)) + geom_point() + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_text(aes(label = description), nudge_y = 0.1) + 
  geom_linerange(aes(xmin = lb2, xmax = ub2)) + 
  labs(y = "", x = "Risk difference (percentage points)")

#write_csv(results4, file = "/home/corinneriddell/ACA-PTB/output-data/results.csv")

```

### Dynamic TWFE specifications

Here the exposure is coded as time since the policy changes. This is equal to 0 for never treated states (all their state months), and 0 in pre-treatment
time for treated states. And then it counts the time since the policy change for each following month.

```{r}
mod7<- glm(ptb_prop ~ as.factor(time_since_policy) + state_r + as.factor(month_index), data = all_years_agg, family = "gaussian") 
#Gaussian to keep things additive
summary(mod7)

#robust SE
system.time(m7_var <- vcovHC(mod7, type="HC3"))

dynamic_effects <- data.frame("event_time" = NA,
                              "estimate" = NA,
                              "lb" = NA,
                              "ub" = NA)

for (i in 1:36) {
  estimate <- summary(mod7)$coefficients[paste0("as.factor(time_since_policy)", i), "Estimate"]
  me <- 1.96*sqrt(m7_var[paste0("as.factor(time_since_policy)", i), paste0("as.factor(time_since_policy)", i)])
  new_row <- data.frame("event_time" = i, "estimate" = estimate, "lb"= estimate - me, "ub" = estimate + me)
  dynamic_effects[i, ] <- c(i, estimate, estimate - me, estimate + me)
}


ggplot(dynamic_effects, aes(x = event_time, y = estimate)) + 
  geom_point() + 
  geom_linerange(aes(ymin = lb, ymax = ub )) + 
  geom_hline(yintercept = 0)
```