---
title: "Callaway & Sant'Anna and Sun & Abraham Estimates"
output: html_document
---

**Overview**

In this file we run the Callaway & Sant'Anna and Sun & Abraham approaches. We also run the bacon decomposition.

**Inputs:**

Note that the restricted births data is not available publicly and so is not uploaded to GitHub.
./restricted-data/births_all: this is the dataset of individual birth records merged to policy data
./ACA-PTB/output-data/results.csv: These are the results from the TWFE specifications

```{r load-libraries, echo=FALSE, message=F}
# set path
riddellPath <- c( "/data/riddell//R/x86_64-pc-linux-gnu-library/4.1/" , .libPaths() )
.libPaths(riddellPath)

# load libraries
library(magrittr)
library(tidyverse)
library(tibble)
library(readxl)
library(readr)
library(data.table)
library(ggrepel)
library(sandwich)
library(tigris)
library(did)
library(staggered)
library(bacondecomp)
library(TwoWayFEWeights)
```

```{r}
results4 <- read_csv(file = "/home/corinneriddell/ACA-PTB/output-data/results.csv")
```

```{r, echo=FALSE, message=F}
all_years2 <- fread("./restricted-data/births_all")
```

```{r, echo=FALSE, message=F}
#aggregated the individual level data into state-months
all_years_agg <- all_years2 %>% 
  group_by(state_r, month_index) %>% 
  summarize(ptb_prop = mean(CR_ptb), 
            DOB_MM = first(DOB_MM), 
            DOB_YY = first(DOB_YY), 
            post_policy = first(post_policy), 
            ever_treated = first(ever_treated), 
            Expansion_Date = first(Expansion_Date)) %>%
  arrange(state_r, month_index)
```

```{r, echo=FALSE, message=F}
# add variable called treated_group for use in the att_ge() function that indicates the month index of 
# the time of the policy change. It is equal to 0 for states that did not change
all_years_agg <- all_years_agg %>% 
  mutate(treated_group = case_when(Expansion_Date == as.Date("2014-01-01") ~ 37, #16 states (36th month + 1)
                                   Expansion_Date == as.Date("2015-01-01") ~ 49, #PA only
                                   Expansion_Date == as.Date("2015-09-01") ~ 57, #AK only
                                   Expansion_Date == as.Date("2016-07-01") ~ 67, #LA only
                                   Expansion_Date == as.Date("2019-01-01") ~ 0,
                                   Expansion_Date == as.Date("2020-01-01") ~ 0, #late expanders are considered not treated for this analysis since they expanded after 2017
                                   is.na(Expansion_Date) ~ 0) #17 states are not treated (across these last two conditions)
  )

# add another variable called treated_group2 where the never-treated states have a value of Inf rather than 0
# for use in the Roth and Sant'Anna `staggered()` function
all_years_agg$treated_group2 <- all_years_agg$treated_group
all_years_agg$treated_group2[all_years_agg$treated_group == 0] <- Inf
table(all_years_agg$treated_group, all_years_agg$treated_group2, useNA = "always")

#check the coding
# table(all_years_agg$Expansion_Date)
# table(all_years_agg$state_r, all_years_agg$Expansion_Date, useNA = "always")
# 
# table(all_years_agg$treated_group, useNA = "always")

#all_years_agg %>% filter(is.na(treated_group)) %>% View()
```

```{r}
# add numeric state codes for use in the later functions

fips <- fips_codes %>% select(state, state_code, state_name) %>% distinct() %>% filter(state_code < 60) %>% rename(state_r = state)

all_years_agg <- merge(all_years_agg, fips, by = "state_r")

all_years_agg <- all_years_agg %>%  mutate(state_num = as.numeric(state_code))
```


```{r}
#In a new dataset, take away the states that make things unbalanced (balance = the same months of data for each state in the dataset)
#some of the functions require balanced data so we make another dataset that excludes states that lead to imbalance
all_years_agg2 <- all_years_agg %>% filter(! (state_r %in% c("AK", "MA", "MN", "MS", "VA")))

table(all_years_agg2$state_r, all_years_agg2$month_index) #balanced panel
```

### Callaway and Sant'Anna

Ref = never treated
Imbalance is permitted

```{r}
# estimate effects using group-time ATT

m2 <- att_gt(yname = "ptb_prop", 
             tname = "month_index", 
             idname = "state_num", 
             gname = "treated_group", 
             allow_unbalanced_panel = TRUE,
             control_group = "nevertreated",
             data = all_years_agg, 
             anticipation = 0)

m2_ag <- aggte(m2, type="simple", na.rm = T)

m2_ag

m2_ag$overall.att
m2_ag$overall.se

lb = m2_ag$overall.att - 1.96*m2_ag$overall.se 
ub = m2_ag$overall.att + 1.96*m2_ag$overall.se

round(c(m2_ag$overall.att, lb, ub)*100, 2)

#add these estimates to the dataset with the TWFE estimates
results5 <- data.frame(estimate = m2_ag$overall.att,
                       lb = lb, 
                       ub = ub, 
                       method = "Callaway Sant'Anna unbalanced, ref = never treated"
                       )

results5 <- results5 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results5 <- bind_rows(results5, results4)
```

Below we change the reference group:

Ref = not yet treated 
Imbalance is allowed

```{r}
m3 <- att_gt(yname = "ptb_prop", 
             tname = "month_index", 
             idname = "state_num", 
             gname = "treated_group", 
             allow_unbalanced_panel = TRUE,
             control_group = "notyettreated",
             data = all_years_agg, 
             anticipation = 0)

m3_ag <- aggte(m3, type="simple", na.rm = T)

m3_ag

m3_ag$overall.att
m3_ag$overall.se

lb = m3_ag$overall.att - 1.96*m3_ag$overall.se 
ub = m3_ag$overall.att + 1.96*m3_ag$overall.se

round(c(m3_ag$overall.att, lb, ub)*100, 2)

results6 <- data.frame(estimate = m3_ag$overall.att,
                       lb = lb, 
                       ub = ub, 
                       method = "Callaway Sant'Anna unbalanced, ref = not yet treated"
                       )

results6 <- results6 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results6 <- bind_rows(results6, results5)
```

The warning "Warning in pre_process_did(yname = yname, tname = tname, idname = idname,  :
  Be aware that there are some small groups in your dataset.
  Check groups: 49,57,67."
  
is because Pennsylvania, Alaska, and Louisiana introduced ACA expansion at t=49, 57, and 67 at time points when no other states introduced the change. So estimating the effect for these groups individually will be imprecise.

The other warnings "No units in group 57 in time period X" warn that Alaska was introduced to the panel late because of inclusion criteria.

Here we re-run the model but on `all_years_agg2`, the dataset we created that is balanced.

```{r}
m2_2 <- att_gt(yname = "ptb_prop", 
             tname = "month_index", 
             idname = "state_num", 
             gname = "treated_group", 
             allow_unbalanced_panel = FALSE,
             control_group = "nevertreated",
             data = all_years_agg2, 
             anticipation = 0)

m2_ag_2 <- aggte(m2_2, type="simple")

m2_ag_2

m2_by_group_2 <- aggte(m2_2, type = "group")
summary(m2_by_group_2)

m2_ag_2$overall.att
m2_ag_2$overall.se

lb_2 = m2_ag_2$overall.att - 1.96*m2_ag_2$overall.se 
ub_2 = m2_ag_2$overall.att + 1.96*m2_ag_2$overall.se

round(c(m2_ag_2$overall.att, lb_2, ub_2)*100, 2)

results7 <- data.frame(estimate = m2_ag_2$overall.att,
                       lb = lb_2, 
                       ub = ub_2, 
                       method = "Callaway Sant'Anna manual balance, ref = never treated"
                       )

results7 <- results7 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results7 <- bind_rows(results7, results6)
```

Here we rerun the model but restricted to states that implementing the policy change (ever-treated only)

```{r}
# estimate effects using group-time ATT for only those who are not yet treated
m2_ea <- att_gt(yname="ptb_prop", 
                tname="month_index", 
                idname="state_num",
                gname="treated_group",
                allow_unbalanced_panel = FALSE,
                control_group = "notyettreated",
                data=all_years_agg %>% filter(ever_treated == T), 
                anticipation=0)

m2_ea_ag <- aggte(m2_ea, type="simple")

m2_ea_ag$overall.att
m2_ea_ag$overall.se

lb = m2_ea_ag$overall.att - 1.96*m2_ea_ag$overall.se 
ub = m2_ea_ag$overall.att + 1.96*m2_ea_ag$overall.se
round(c(m2_ea_ag$overall.att, lb, ub)*100, 2)

results8 <- data.frame(estimate = m2_ea_ag$overall.att,
                       lb = lb, 
                       ub = ub, 
                       method = "Callaway Sant'Anna ever-treated only, ref = notyettreated"
                       )

results8 <- results8 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results8 <- bind_rows(results8, results7)
```

### Sun and Abraham method estimate of the simple weighted average 

-This method requires that the data be balanced, and drops states without balance
-This method also drops states with a single cross sectional unit, that is treatment cohorts g = 49 and 67 (PA and LA) 

```{r}
mod7 <- staggered_sa(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  estimand = "simple")
mod7

mod7_event_study2 <- staggered_sa(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  eventTime = -30:35,
                  estimand = "eventstudy")
mod7_event_study2 %>% View()

mod7_event_study2 %<>% mutate(lb = estimate - 1.96*se, ub = estimate + 1.96*se)

ggplot(mod7_event_study2, aes(x = eventTime, y = estimate)) +
  geom_point() + 
  geom_linerange(aes(ymin = lb, ymax = ub)) + theme_bw() + geom_hline(aes(yintercept = 0))

round(c(mod7$estimate, mod7$estimate - 1.96*mod7$se, mod7$estimate + 1.96*mod7$se)*100, 2)
round(c(mod7$estimate, mod7$estimate - 1.96*mod7$se_neyman, mod7$estimate + 1.96*mod7$se_neyman)*100, 2)

results9 <- data.frame(estimate = mod7$estimate,
                       lb = mod7$estimate - 1.96*mod7$se_neyman, 
                       ub = mod7$estimate + 1.96*mod7$se_neyman, 
                       method = "Sun and Abraham, balanced, no small groups"
                       )

results9 <- results9 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results9 <- bind_rows(results9, results8)
```

Callaway Sant'Anna manual balance, ref = never treated, no small groups (to make similar to Sun and Abraham for comparison)

```{r}
mod8 <- att_gt(yname = "ptb_prop", 
             tname = "month_index", 
             idname = "state_num", 
             gname = "treated_group", 
             allow_unbalanced_panel = FALSE,
             control_group = "nevertreated",
             data = all_years_agg2 %>% filter(! (state_r %in% c("LA", "PA"))), 
             anticipation = 0)

mod8_aggte <- aggte(mod8, type="simple")

mod8_aggte$overall.att
mod8_aggte$overall.se

lb_2 = mod8_aggte$overall.att - 1.96*mod8_aggte$overall.se 
ub_2 = mod8_aggte$overall.att + 1.96*mod8_aggte$overall.se

round(c(mod8_aggte$overall.att, lb_2, ub_2)*100, 2)

results10 <- data.frame(estimate = m2_ag_2$overall.att,
                       lb = lb_2, 
                       ub = ub_2, 
                       method = "Callaway Sant'Anna manual balance, no small groups, ref = never treated"
                       )

results10 <- results10 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results10 <- bind_rows(results10, results9)
```

Re-run main analysis but restricted to no small groups and make it balanced (to make similar and comparable to Sun and Abraham)

```{r}
#take away the states that make things unbalanced
all_years_agg3 <- all_years_agg2 %>% filter(! (state_r %in% c("PA", "LA")))

mod6_balanced_restricted <- glm(ptb_prop ~ post_policy + state_r + as.factor(month_index), data = all_years_agg3, family = "gaussian") 
#Gaussian to keep things additive
summary(mod6_balanced_restricted)

#how does this estimate compare to model 2?

#robust SE
system.time(m6_var <- vcovHC(mod6_balanced_restricted, type="HC3"))

lb2 = summary(mod6_balanced_restricted)$coefficients["post_policy", "Estimate"] - 1.96*sqrt(m6_var["post_policy","post_policy"]) 
ub2 = summary(mod6_balanced_restricted)$coefficients["post_policy", "Estimate"] + 1.96*sqrt(m6_var["post_policy","post_policy"])
c(lb2, ub2)


results11 <- data.frame(estimate = summary(mod6_balanced_restricted)$coefficients["post_policy", "Estimate"], 
           lb = lb2, 
           ub = ub2, 
           method = "state-level TWFE LPM balanced restricted")

results11 <- results11 %>% mutate(estimate2 = estimate*100, 
                     lb2 = lb*100,
                     ub2 = ub*100,
                     description = paste0(round(estimate2, 2), " (", round(lb2, 2), ", ", round(ub2, 2), ")"))

results11 <- bind_rows(results11, results10)

```

Investigate the difference between CS and Bacon for Group 37:

```{r}
m2_by_group <- aggte(m2, type = "group")
summary(m2_by_group)
#-0.0002 for 37
#-0.0039 for 49
#-0.0040 for 67 

m2_by_time <- aggte(m2, type = "dynamic", na.rm = T)
summary(m2_by_time)
ggdid(m2_by_time) + geom_hline(yintercept = 0)

point_ests <- m2_by_time$att.egt
lbs <- point_ests - 2.95*m2_by_time$se.egt
ubs <- point_ests + 2.95*m2_by_time$se.egt
event_time <- m2_by_time$egt

#not yet treated
m3_by_time <- aggte(m3, type = "dynamic", na.rm = T)
summary(m3_by_time)
ggdid(m3_by_time) + geom_hline(yintercept = 0)

point_ests_notyet <- m3_by_time$att.egt
lbs_notyet <- point_ests_notyet - 2.95*m3_by_time$se.egt
ubs_notyet <- point_ests_notyet + 2.95*m3_by_time$se.egt


cs_dynamic <- data.frame(event_time, 
                         point_ests, lbs, ubs,
                         point_ests_notyet, lbs_notyet, ubs_notyet
                         )

ggplot(cs_dynamic, aes(x = time, y = point_ests)) + geom_point(aes(col = time >= 0)) + 
  geom_hline(yintercept = 0, lty = 2) + 
  geom_linerange(aes(ymin = lbs, ymax = ubs, col = time >= 0)) + 
  labs(y = "Estimate", x = "Event time") + 
  theme_bw()
```

```{r}
# results6 <- results5 %>% mutate(method2 = factor(method, 
#                                                  levels = c("individual-level TWFE LPM", "state-level TWFE LPM",
#                                                             "state-level TWFE ever-treated", "Callaway Sant'Anna",
#                                                             "Sun and Abraham")))

results11$ever_treated <- 0
results11$ever_treated[results11$method == "state-level TWFE ever-treated" | results11$method ==  "Callaway Sant'Anna ever-treated only, ref = notyettreated"] <- 1

ggplot(results11 %>% filter(ever_treated == 0), aes(y = method, x = estimate2)) + geom_point() + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_text(aes(label = description), nudge_y = 0.1) + 
  geom_linerange(aes(xmin = lb2, xmax = ub2)) + 
  labs(y = "", x = "Risk difference (percentage points)") + 
  scale_y_discrete(limits=rev)
```

Figure 1 

```{r}

results12 <- results11 %>% filter(method %in% c("state-level TWFE LPM", 
                                          "Callaway Sant'Anna unbalanced, ref = never treated",
                                          "Sun and Abraham, balanced, no small groups")) %>%
  mutate(method2 = factor(method, levels = c("state-level TWFE LPM",
                                                "Callaway Sant'Anna unbalanced, ref = never treated",
                                                "Sun and Abraham, balanced, no small groups")))

#remake figure for the paper
ggplot(results12,
       aes(y = method2, x = estimate2)) + geom_point() + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_text(aes(label = description), nudge_y = 0.1) + 
  geom_linerange(aes(xmin = lb2, xmax = ub2)) + 
  labs(y = "", x = "Risk difference (percentage points)") + 
  scale_y_discrete(limits = rev, labels = c("Sun and Abraham", "Callaway and Sant'Anna", "TWFE")) + 
  theme_bw()

```

Figure S1

```{r}
results13 <- results11 %>% filter(method %in% c("state-level TWFE LPM balanced restricted", 
                                          "Callaway Sant'Anna manual balance, no small groups, ref = never treated",
                                          "Sun and Abraham, balanced, no small groups")) %>%
  mutate(method2 = factor(method, levels = c("state-level TWFE LPM balanced restricted",
                                                "Callaway Sant'Anna manual balance, no small groups, ref = never treated",
                                                "Sun and Abraham, balanced, no small groups")))

ggplot(results13,
       aes(y = method2, x = estimate2)) + geom_point() + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_text(aes(label = description), nudge_y = 0.1) + 
  geom_linerange(aes(xmin = lb2, xmax = ub2)) + 
  labs(y = "", x = "Risk difference (percentage points)") + 
  #scale_y_discrete(limits=rev) +
  scale_y_discrete(limits = rev, labels = c("Sun and Abraham", "Callaway and Sant'Anna", "TWFE")) + #"Sun and Abraham", "Callaway and Sant'Anna", "TWFE")) + 
  theme_bw()
```

### Bacon decomposition

```{r}
# df_bacon <- bacon(ptb_prop ~ post_policy,
#                   data = all_years_agg,
#                   id_var = "state_num",
#                   time_var = "month_index")


table(all_years_agg$state_r, all_years_agg$month_index) #states AK, MA, MN, MS, VA exclude the earliest years

# all_years_agg %>% filter(state_r == "AK") %>% View() #treated in month 57 (one of the ones treated on their own)
# all_years_agg %>% filter(state_r == "MA") %>% View() #treated in month 37 along with 13 other states
# all_years_agg %>% filter(state_r == "MN") %>% View() #treated in month 37 along with 13 other states
# all_years_agg %>% filter(state_r == "MS") %>% View() #never-treated
# all_years_agg %>% filter(state_r == "VA") %>% View() #never-treated



df_bacon <- bacon(ptb_prop ~ post_policy,
                  data = all_years_agg2,
                  id_var = "state_num",
                  time_var = "month_index")

(0.06798*0.00074) + (0.02055*-0.00197) + (0.91146*0.00154)
#0.00141347
#91% of the weight is on the treated vs. the untreated and it's average effect is 0.00154

df_bacon

#compare with the results from the model:
#0.0014171 (a touch different but very close)
mod1 <- lm(ptb_prop ~ post_policy + state_r + as.factor(month_index), data = all_years_agg2) 
summary(mod1)
```


```{r}
chaise <- twowayfeweights(all_years_agg2, "ptb_prop", G = "treated_group", T = "month_index", D = "post_policy", cmd_type = "feTR")

chaise2 <- twowayfeweights(all_years_agg2 %>% filter(month_index >= 33, ever_treated == T), "ptb_prop", G = "treated_group", T = "month_index", D = "post_policy", cmd_type = "feTR")

df_bacon2 <- bacon(ptb_prop ~ post_policy,
                  data = all_years_agg2 %>% filter(month_index >= 33, ever_treated == T),
                  id_var = "state_num",
                  time_var = "month_index")

df_bacon2
```

Related to prob treatment at any time point, 
If half are never treated the half ever treated is 0.5

### Roth and Sant'Anna

I don't think we want to include these but keeping them here in case we do...

```{r}

#Calculate efficient estimator for the simple weighted average
mod3 <- staggered(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  estimand = "simple")

mod3

round(c(mod3$estimate, mod3$estimate - 1.96*mod3$se, mod3$estimate + 1.96*mod3$se)*100, 2)
round(c(mod3$estimate, mod3$estimate - 1.96*mod3$se_neyman, mod3$estimate + 1.96*mod3$se_neyman)*100, 2)

mod3_1 <- staggered(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  estimand = "cohort")

mod3_1

round(c(mod3_1$estimate, mod3_1$estimate - 1.96*mod3_1$se, mod3_1$estimate + 1.96*mod3_1$se)*100, 2)
round(c(mod3_1$estimate, mod3_1$estimate - 1.96*mod3_1$se_neyman, mod3_1$estimate + 1.96*mod3_1$se_neyman)*100, 2)


mod3_2 <- staggered(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  estimand = "calendar")

mod3_2

round(c(mod3_2$estimate, mod3_2$estimate - 1.96*mod3_2$se, mod3_2$estimate + 1.96*mod3_2$se)*100, 2)
round(c(mod3_2$estimate, mod3_2$estimate - 1.96*mod3_2$se_neyman, mod3_2$estimate + 1.96*mod3_2$se_neyman)*100, 2)
```


### OLD

```{r}
# Reproduce the Callaway and Sant'Anna estimates using the staggered package

#example
#df <- staggered::pj_officer_level_balanced #load the officer data
# staggered(df = df, 
#           i = "uid",
#           t = "period",
#           g = "first_trained",
#           y = "complaints", 
#           estimand = "simple")

#Callaway and Sant'Anna - reproduce previous findings (with 49 and 67 dropped, curious that 57 is not also dropped...)
mod1 <- staggered_cs(df = all_years_agg,
                  i = "state_num",
                  t = "month_index",
                  g = "treated_group2",
                  y = "ptb_prop", 
                  estimand = "simple")
mod1

round(c(mod1$estimate, mod1$estimate - 1.96*mod1$se, mod1$estimate + 1.96*mod1$se)*100, 2)
round(c(mod1$estimate, mod1$estimate - 1.96*mod1$se_neyman, mod1$estimate + 1.96*mod1$se_neyman)*100, 2)
```

```{r}
# estimate effects using group-time ATT
a2 <- all_years_agg %>% filter(! (treated_group %in% c(49, 67)))

m2 <- att_gt(yname = "ptb_prop", 
             tname = "month_index", 
             idname = "state_num", 
             gname = "treated_group", 
             data = a2, 
             anticipation = 0)

m2_ag <- aggte(m2, type="simple")
m2_ag$overall.att
m2_ag$overall.se

lb = m2_ag$overall.att - 1.96*m2_ag$overall.se 
ub = m2_ag$overall.att + 1.96*m2_ag$overall.se
round(c(m2_ag$overall.att, lb, ub)*100, 2)
```
