---
title: "Companion Guide to 'New developments in difference-in-differences designs'"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

## Overview

In this companion guide, we illustrate scenarios in which difference in differences
analysis can be applied. We start with the simplest of scenarios and slowly amp
up the complexity. For each scenario we describe a target parameter
to estimate and the parameter estimated by a two-way fixed effects (TWFE) model. We
also apply the Goodman-Bacon decomposition to this parameter to determine if the
TWFE estimate is partially composed of estimates that are "forbidden" (e.g., ones
that compare a newly treated state to a previously treated state) and apply alternate
estimation approaches. 

The goal is to illustrate scenarios where the usual TWFE method of estimation 
provides suitable results and scenarios where this approach is biased. 
In the latter case, we show alternative methods to estimation to overcome the 
issue.

In the following examples, state policy changes are the effects of interest. 

### Load the R packages used in this analysis

First, load the packages for reading and plotting the data. The packages `bacondecomp` and 
`did` are specific to DID analyses.

```{r load-libraries, warning=F, message=F}
# you will need to install these packages if you don't have them already
# install.packages(c("here", "readxl", "tidyverse", "patchwork", "magrittr", 
# "broom", "ggrepel",  "bacondecomp", "did"))

library(here)
library(readxl)
library(tidyverse)
library(patchwork)
library(magrittr)
library(broom)
library(ggrepel)
library(bacondecomp)
library(did)
```

## Scenario 1

We first consider the simplest case in which there is one state that never adopts 
the policy (the never-treated group), and one state that does implement the policy 
(the group that undergoes treatment). The groups are observed during two time 
periods only. 

Below, the data are read in from an excel spreadsheet. Feel free to view the 
data in the spreadsheet or use an R `View()`er window.
The data contain four variables `state`, `time`, `policy`, and `outcome`. 

* `state`: ID for the grouping variable
* `time`: Time index, begins at 1
* `policy`: Indicator variable for exposure to the policy change
* `outcome`: The outcome variable

```{r}
s1 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen1",
                col_types = c("text", "text", "text", "numeric"))
# str(s1)
# View(s1)
```
Note that I specified the column types for each variable. I read in the state, 
time, and policy variables as "text" so that R will consider them as 
categorical/factor variables in the analysis (or, using econ-speak using 
"fixed effects").

### Visualization of Scenario 1

```{r, echo = F}
ggplot(s1, aes(x = time, y = outcome)) + 
  geom_line(aes(group = state, col = state)) +
  geom_point(aes(fill = state, pch = state), size = 4) +
  geom_text(aes(label = outcome), nudge_x = 0.075) +
  scale_fill_manual(values = c("#a6cee3","#005a32")) + 
  scale_color_manual(values = c("#a6cee3","#005a32")) +
  scale_shape_manual(values = c(21, 23)) + 
  theme_bw()
```

The following can be read from the labelled plot:

* Pre-treatment difference of 1
* Post-treatment difference of 3
* Causal effect of policy = 3-1 = 2 
or equivalently, 
* State 1 difference of 5
* State 2 difference of 3
* Causal effect of policy = 5-3 = 2 
under the following assumptions

Three standard assumptions are typical stated in DID papers and required for the 
DID estimate [in the simple case?] to estimate the causal effect of interest:

1) **Parallel trends, or counterfactual parallel trends as described by 
Ben-Michael:** Had the treated group not been treated, the trends for the treated 
state would be the same as the trends for the never-treated state.
2) **No anticipation:** The effect of the policy change begins after it is introduced; 
there is no anticipatory effect before the policy is introduced. Anticipatory
effects can occur if individuals change their behavior in anticipation of a 
policy change.
3) **[NAME FOR THIS ONE:]** The policy is the only thing that changes at the time of
treatment that affects the outcome

If you're familiar with causal inference, you will also know about the main 
assumptions required when estimating causal effects. Ben-Micheal et al. expound
on these in their paper, and we expand on them here:

4) **Consistency:** The policy change is the same across treated units. For 
example, if states introduced a texting ban while driving and these bans were 
associated with different penalties, then the policy change would not meet the
consistency assumption because the different penalties may have varying effects 
on the outcomes of interest.
5) **No interference:** Outcomes of individuals in one state do not affect 
outcomes of individuals in other states.
6) **Exchangeability:** Had the treated states been untreated, they would have 
experienced the trend in outcomes experienced by the untreated group. This 
causal assumption is the same as the counterfactual parallel trends assumption
[DANA DO YOU AGREE?]. 
7) **Positivity:** Each state has a non-zero probability of being treated. [TO
DISCUSS: See Ben-Micheal paper - they say that positivity is not a required
assumption for standard DID]

WHERE TO PUT THIS ONE IF WE TALK ABOUT IT?

8) **Correct model form**: SOMETHING ABOUT LINEARITY? ALSO RELATED TO THE PARALLEL
TRENDS ASSUMPTION 


### Two-way fixed effected (TWFE) regression model

Under the canonical TWFE design, we can estimate the policy effect by including
a fixed effect (FE) for state, a FE for time, and an indicator for the policy
change. The effect estimate is the regression coefficient on the policy variable.

```{r}
s1_mod <- lm(outcome ~ state + time + policy, data = s1)
tidy(s1_mod)
```

The coefficient on the policy term is `r s1_mod$coefficients[4]` showing that
the regression model estimate equalled the causal effect.

**Takeaway:** When there are only two treatment groups and two time points, 
where one of the groups becomes treated, you can calculate the DID estimate 
by hand or using TWFE regression. Note that there are only 4 rows of data
and no sampling variability so no measures of variation can be estimated in this
setting.

## Scenario 2A and 2B

This scenario is similar to Scenario 1, except now there are two states that
undergo treatment. In Scenario 2A, the magnitude of the treatment effect is the 
same in both states. In Scenrio 2B, the magnitude of the treatment effect 
varies by state.

```{r import-scen2}
s2a <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen2a",  
                 col_types = c("text", "text", "text", "numeric"))

s2b <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen2b",
                  col_types = c("text", "text", "text", "numeric"))

s2a %<>% mutate(ever_trt = case_when(state %in% c("2", "3") ~ "treated",
                                     state == "1" ~ "never-treated"))

s2b %<>% mutate(ever_trt = case_when(state %in% c("2", "3") ~ "treated",
                                     state == "1" ~ "never-treated"))
```

### Visualization of Scenario 2

```{r, echo=F}
s2a_plot <-ggplot(s2a, aes(x = time, y = outcome)) + 
  geom_line(aes(group = state, col = state)) +
  geom_point(aes(fill = state, pch = ever_trt), size = 4) +
  geom_text(aes(label = outcome), nudge_x = 0.1) +
  scale_fill_manual(values = c("#a6cee3","#005a32", "#99000d"), guide = "none") +
  scale_color_manual(values = c("#a6cee3","#005a32", "#99000d")) +
  scale_shape_manual(values = c(21, 23)) + 
  theme_bw() + labs(title = "Scenario 2a")

s2b_plot <-ggplot(s2b, aes(x = time, y = outcome)) + 
  geom_line(aes(group = state, col = state)) +
  geom_point(aes(fill = state, pch = ever_trt), size = 4) +
  geom_text(aes(label = outcome), nudge_x = 0.1) +
  scale_fill_manual(values = c("#a6cee3","#005a32", "#99000d"), guide = "none") +
  scale_color_manual(values = c("#a6cee3","#005a32", "#99000d")) +
  scale_shape_manual(values = c(21, 23)) + 
  theme_bw() + labs(title = "Scenario 2b")

s2a_plot + s2b_plot + plot_layout(guides = "collect")  & 
  scale_y_continuous(limits = range(c(s2a$outcome, s2b$outcome)))
```

The following can be read from the labelled plot for Scenario 2A:

* Post-Pre Difference for never-treated state 1: 12-9=3
* Post-Pre Difference for treated state 2: 15-10=5
* Post-Pre Difference for treated state 3: 16-11=5
* Causal effect of policy = 5-3 for both states for an average treatment effect on the treated (ATT) of 2

The following can be read from the labelled plot for Scenario 2B:

* Post-Pre Difference for never-treated state 1: 12-9=3
* Post-Pre Difference for treated state 2: 15-10=5
* Post-Pre Difference for treated state 3: 17-11=6
* Causal effect of policy = 2 for state 2 vs 1 and 3 for state 3 vs 1 for an ATT of 2.5

## TWFE Regression model

```{r}
s2a_mod <- lm(outcome ~ policy + state + time, data = s2a)
tidy(s2a_mod)
```

The coefficient estimate equals to `r s2a_mod$coefficients["policy1"]` the ATT for Scenario 2a. 

```{r}
s2b_mod <- lm(outcome ~ policy + state + time, data = s2b)
tidy(s2b_mod)
```

The coefficient estimate equals to `r s2b_mod$coefficients["policy1"]` the ATT for Scenario 2b. 

**Takeaway:** When there are multiple treatment groups and two time points, 
where >1 of the groups becomes treated, you can calculate the DID estimate by 
hand or using TWFE regression. If treatment effects are heterogeneous across 
states, then the estimated effect is the average ATT across the 
states

**Things to discuss with Dana:**

- The state FEs are a bit off (FE for state 3 should be 2, for state 2 should be 1)
- Interesting to think about differences in the p-value for scen 2a (very small) vs 2b (big)

## Scenario 3A and 3B

In scenario 3, the number of time periods is increased to incorporate 5 time 
points before and after a policy change. One never-treated and one treated
group are considered. In Scenario 3A, the effect of treatment is constant once
treatment is introduced. In Scenario 3B, the effect of treatment increases with 
time (i.e., it is dynamic).

Because there are multiple time points, when we read in the data we update the 
column type for the time variable to be numeric. This will help with plotting 
the data. Later, when we run the TWFE model, we will use `factor(time)` to
model time using indicator variables (i.e., time fixed effects) rather than
including time as a linear term.

```{r}
s3a <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen3a",
                 col_types = c("text", "numeric", "text", "numeric", "text"))

s3b <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen3b",
                 col_types = c("text", "numeric", "text", "numeric", "text"))
```

In addition to `state`, `time`, `policy`, and `outcome`, the data contains 
another variable `time_since_policy` which equals 0 before treatment (and 
always equals 0 for the never-treated), and counts the time since treatment in
the treated state.

```{r, echo = F}

# Because there are multiple time points, we first take the average of the outcome
# during the pre and post-treatment time period separately for each state. 

#first, add an indicator indicating time post treatment.
s3a %<>% mutate(post = time >= 6) 
s3b %<>% mutate(post = time >= 6)

s3a_mean_info <- s3a %>% 
  group_by(state, post) %>%
  summarise(mean_y = mean(outcome)) %>% 
  pivot_wider(id_cols = state, 
              names_from = post, 
              names_pre = "post_",
              values_from = mean_y) %>%
  mutate(diff = post_TRUE - post_FALSE)

s3b_mean_info <- s3b %>% group_by(state, post) %>%
  summarise(mean_y = mean(outcome)) %>% 
  pivot_wider(id_cols = state,
              names_from = post,
              names_pre = "post_",
              values_from = mean_y) %>%
  mutate(diff = post_TRUE - post_FALSE)
```
## Visualization of Scenario 3

```{r, echo=F}

s3a_plot <- ggplot(s3a, aes(x = time, y = outcome)) + 
  geom_line(aes(col = state)) + 
  geom_point(aes(fill = state, pch = state)) +
  labs(title = "Scenario 3A (Constant change)") +
  geom_vline(aes(xintercept = 5.5), lty = 2) + 
  geom_label(data = s3a_mean_info, 
            aes(y = post_FALSE, x = 4, 
                col = state, 
                label = paste("y_mean= ", post_FALSE)),
            nudge_y = 0.1) +
    geom_label(data = s3a_mean_info, 
            aes(y = post_TRUE, x = 8, 
                col = state, 
                label = paste("y_mean= ", post_TRUE)),
            nudge_y = 0.1) +
  scale_fill_manual(values = c("#a6cee3","#005a32")) +
  scale_color_manual(values = c("#a6cee3","#005a32")) +
  scale_shape_manual(values = c(21, 23)) + 
  theme_bw()

s3b_plot <- ggplot(s3b, aes(x = time, y = outcome)) + 
  geom_line(aes(col = state)) + 
  geom_point(aes(fill = state), pch = 21) + 
  labs(title = "Scenario 3B (Dynamic change)") +
  geom_vline(aes(xintercept = 5.5), lty = 2) + 
  geom_label(data = s3b_mean_info, 
            aes(y = post_FALSE, x = 4, 
                col = state, 
                label = paste("y_mean= ", post_FALSE)),
            nudge_y = 0.1) +
    geom_label(data = s3b_mean_info, 
            aes(y = post_TRUE, x = 8, 
                col = state, 
                label = paste("y_mean= ", post_TRUE)),
            nudge_y = 0.1) +
  scale_fill_manual(values = c("#a6cee3","#005a32")) +
  scale_color_manual(values = c("#a6cee3","#005a32")) +
  scale_shape_manual(values = c(21, 23)) + 
  theme_bw()

s3a_plot + s3b_plot + plot_layout(guides = "collect")  & 
  scale_y_continuous(limits = range(c(s3a$outcome, s3b$outcome)))
```

The following can be read from the labelled plot for Scenario 3A:

* Pre-treatment difference is 17-15=2
* Post-treatment difference is 34-30=4
* DID = 2

The following can be read from the labelled plot for Scenario 3B:

* Pre-treatment difference is 17-15=2
* Post-treatment difference is 38-30=8
* DID = 6

For Scenario 3B the effect is dynamic, and rather than 
calculating an ATT that aggregates over post-time, it might be beneficial to calculate 
the effect separately by time since treatment. Here, the effect increases over 
time, suggesting that the policy change takes time to "kick-in". In contrast, 
some policies may be associated with large initial effects that attenuate over 
time. Knowing if a policy introduction is associated with a changing effect over
time is therefore important for understanding the real-world effects of the policy. 

Thus, in this scenario we first estimate the average treatment effect using a 
TWFE model, and then we estimate the dynamic effect by extending the TWFE model
to include a categorical indicator for time since the policy's introduction. 

## TWFE Regression Model

Estimation of the dynamic effect in Scenario 3A:
```{r}
s3a_mod <- lm(outcome ~ policy + state + factor(time), 
             data = s3a)
tidy(s3a_mod)
```

The coefficient of the policy term is `r s3a_mod$coefficients["policy1"]`
showing that the regression model estimate equaled the causal effect.

Estimation of the dynamic effect in Scenario 3B:
```{r}
s3b_mod <- lm(outcome ~ policy + state + factor(time), 
             data = s3b)
tidy(s3b_mod)
```

The coefficient of the policy term is `r s3b_mod$coefficients["policy1"]`
showing that the regression model estimate equaled the causal effect.

## TWFE Regression Model with `time_since_policy` specification

Rather than using a binary indicator for the policy change, it can be coded
using the `time_since_policy` variable. If we include this as a factor variable
in the model, then separate effects will be estimated for each time since 
treatment. 

For Scenario 3A, using this variable should yield 2 for each time since treatment,
since the treatment effect is constant over time.

```{r}
s3a_mod2 <- lm(outcome ~ time_since_policy + state + factor(time), 
             data = s3a)
tidy(s3a_mod2)
```

Indeed, this is what we see!

For Scenario 3B, the effect increases by 2 units for every unit of time since
treatment is initiated so this should be reflected in the coefficient estimates.

```{r}
s3b_mod2 <- lm(outcome ~ time_since_policy + state + factor(time), 
             data = s3b)
tidy(s3b_mod2)
```

For this scenario, the effect estimate is 2 in the first time after treatment, 
then 4, and so on. 

Note that we could have modeled `time_since_policy` linearly, however including 
`time_since_policy` as a factor variable allows the estimation of effects to be
non-linear over time.

## Scenario 4

Scenario 4 extends Scenario 3 to the multiple group setting. State 1 is 
never-treated, while states 2-4 undergo treatment at time=6. The effects are 
not dynamic in time, but they are heterogeneous with some states having a larger
treatment effect 

```{r}
s4 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen4",
                col_types = c("text", "numeric", "text", "numeric", "text", "numeric"))
```
(I WOULD CHANGE THE A MARKERS TO CIRCLES)

```{r, echo=F}
s4 %<>% mutate(post = time >= 6)

s4_mean_info <- s4 %>% group_by(state, post) %>%
  summarise(mean_y = mean(outcome)) %>% 
  pivot_wider(id_cols = state, names_from = post, names_pre = "post_", values_from = mean_y) %>%
  mutate(diff = post_TRUE - post_FALSE)
```
## Visualization of Scenario 4

```{r, echo=F}
ggplot(s4, aes(x = time, y = outcome)) + 
  geom_line(aes(col = state)) + 
  geom_point(aes(fill = state), pch = 21) + labs(title = "Scenario 4") +
  geom_vline(aes(xintercept = 5.5), lty = 2) + 
  geom_label(data = s4_mean_info, 
            aes(y = post_FALSE, x = 4, 
                col = state, 
                label = paste("y_mean= ", post_FALSE))) +
    geom_label(data = s4_mean_info, 
            aes(y = post_TRUE, x = 8, 
                col = state, 
                label = paste("y_mean= ", post_TRUE))) + 
  theme_bw() +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) + 
  scale_fill_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d"))
```

The following can be read from the labelled plot:

* Pre-post difference for never-treated state 1: 30-15 = 15
* Pre-post difference for treated state 2: 34-17 = 17
* Pre-post difference for treated state 3: 35-19 = 16
* Pre-post difference for treated state 4: 39-21 = 18

* $DID_{2 vs. 1} = 17-15 = 2$
* $DID_{3 vs. 1} = 16-15 = 1$
* $DID_{4 vs. 1} = 18-15 = 3$
* $ATT = 2$

## TWFE Regression Model 

```{r}
s4_mod <- lm(outcome ~ policy + state + factor(time), 
             data = s4)
tidy(s4_mod)
```

The coefficient of the policy term is `r s4_mod$coefficients["policy1"]`
showing that the regression model estimates the ATT.

One of the issues with TWFE estimation is that contrasts between earlier-treated
and later-treated states contribute to the coefficient estimate of the policy 
term. While all states are treated at the same time in Scenario 4, we can use
the Goodman-Bacon decomposition function `bacon()` to partition the effect 
estimate into weighted component parts. 

Before using the `bacon()` function, we need to modify some of the variables
that are stored as characters by R to be stored as numeric: 

```{r}
s4 %<>% mutate(state_n = as.numeric(as.character(state)),
               policy_n = as.numeric(as.character(policy)))
```

The `bacon` function has an argument called `formula`, where you list the outcome
as a function of the policy change, i.e., `formula = outcome ~ policy_n`. Note 
that no fixed effects are included in the formula. The time and state fixed 
effects are specified by the `time_var`and `id_var` arguments.

```{r}
s4_bacon <- bacon(formula = outcome ~ policy_n, 
                  data = s4,
                  id_var = "state_n",
                  time_var = "time")
```

This small table illustrates that 100% of the weight is put on comparisons between
treated and untreated states and the ATT equals 2.

For more detail we can print the object:

```{r}
s4_bacon 
```

`treated=6` says that all treated states start treatment at time = 6 and are
compared to untreated states (with a fictitious treatment time of 99999).

The results from the Goodman-Bacon Decomposition are reassuring and tell us we
can trust the TWFE estimate. For pedagodgical purposes, we also show the estimate
if we applied the Callaway Sant'Anna estimator using the `att_gt` function to 
estimate group-time ATEs.

Like `bacon()`, `att_gt()` requires all numeric variables. It also requires an
argument called `gname`. `gname` expects a variable which encodes for each state
the time of first policy change. For never-treated state 1, this variable 
equals 0 and for treated states 2-4 this variable equals 6. 

```{r}
s4_cs <- att_gt(yname = "outcome", 
             tname = "time", 
             idname = "state_n", 
             gname = "time_first_treat", 
             data = s4, 
             anticipation = 0)

m2_ag <- aggte(s4_cs, type="simple")
summary(m2_ag)
```

Here, we see the ATT is equal to `r # m2_ag$overall.att` as estimated by the TWFE 
model.

### Scenario 5

Scenario 5 expands upon Scenario 4 by introducing staggered timing. In this 
example, the treated states introduce the policy at two different time points.
The treatment effects are non-dynamic and there are two never-treated states and
two ever-treated states.

(SINCE THIS IS AN EXTENSION OF SCENARIO 5, I WOULD KEEP THE HETEROGENEOUS TX EFFECTS PART
AND JUST INRODUCE STAGGERED TIMING)
```{r}
s5 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen5",
                col_types = c("text", "numeric", "text", "numeric", "text", "numeric"))
```

### Visualization of Scenario 5

```{r, echo = F}
ggplot(s5, aes(x = time, y = outcome)) + 
  geom_line(aes(col = state)) + 
  geom_point(aes(fill = state), pch = 21) + labs(title = "Scenario 5 (Staggered timing)") +
  geom_vline(aes(xintercept = 4.5), lty = 2, col = "#005a32") + 
  geom_vline(aes(xintercept = 11.5), lty = 2, col = "#99000d") +
  scale_fill_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) +
  theme_bw()
```

**Visualization of all comparisons made by TWFE regression**

```{r, echo = F}
mean_1_4 <- s5 %>% filter(time < 5) %>%
  group_by(state) %>%
  summarise(mean_1_4 = mean(outcome))

mean_5_20 <- s5 %>% filter(time >= 5) %>%
  group_by(state) %>%
  summarise(mean_5_20 = mean(outcome))

DID1 <- bind_cols(mean_1_4, 
                  mean_5_20 %>% select(-state)) %>%
  mutate(time_diff1 = mean_5_20 - mean_1_4)

mean_1_11 <- s5 %>% filter(time < 12) %>%
  group_by(state) %>%
  summarise(mean_1_11 = mean(outcome))

mean_12_20 <- s5 %>% filter(time >= 12) %>%
  group_by(state) %>%
  summarise(mean_12_20 = mean(outcome))

DID2 <- bind_cols(mean_1_11, 
                  mean_12_20 %>% select(-state)) %>%
  mutate(time_diff2 = mean_12_20 - mean_1_11)

mean_5_11 <- s5 %>% filter(time >= 5 & time < 12) %>%
  group_by(state) %>%
  summarise(mean_5_11 = mean(outcome))

DID3 <- bind_cols(mean_1_4, 
                  mean_5_11 %>% select(-state)) %>%
  mutate(time_diff3 = mean_5_11 - mean_1_4)

DID4 <- bind_cols(mean_5_11, 
                  mean_12_20 %>% select(-state)) %>%
  mutate(time_diff4 = mean_12_20 - mean_5_11)
```

```{r, echo = F, fig.height = 8, fig.width = 10}
s5_comp1 <- ggplot(s5 %>% filter(state %in% c(1, 2, 3)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "A) Contrast 1 (Earlier treated vs. untreated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32")) +
  theme_bw() + 
  scale_x_continuous(limits = c(1, 20)) + 
  geom_label_repel(data = DID1 %>% filter(state %in% c(1,2,3)), 
                  aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = DID1 %>% filter(state %in% c(1,2,3)), 
                  aes(y = mean_5_20, x = 12.5, label = paste("mean_y:", mean_5_20), col = state))

s5_comp2 <- ggplot(s5 %>% filter(state %in% c(1, 2, 4)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "B) Contrast 2 (Later treated vs. untreated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = DID2 %>% filter(state %in% c(1,2,4)), 
                   aes(y = mean_1_11, x = 6, label = paste("mean_y:", mean_1_11), col = state)) +
  geom_label_repel(data = DID2 %>% filter(state %in% c(1,2,4)), 
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s5_comp4 <- ggplot(s5 %>% filter(state %in% c(3, 4), time >= 5), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "C) Contrast 4 (Earlier vs. later treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = DID4 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state)) +
  geom_label_repel(data = DID4 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s5_comp3 <- ggplot(s5 %>% filter(state %in% c(3, 4), time < 12), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "D) Contrast 3 (Later vs. earlier treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = DID3 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = DID3 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state))

s5_comp1 + s5_comp2 + s5_comp3 + s5_comp4 + plot_layout(guides = "collect")
```

**Contrast 1:**

* Pre-post difference for never-treated state 1 is: 43.5-13.5=30
* Pre-post difference for never-treated state 2 is: 44.5-14.5=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 3 is 49.5-16.5 = 33
* $DID_{3 vs 1,2} = 33-30=3$

**Contrast 2:**

* Pre-post difference for never-treated state 1 is: 54-24=30
* Pre-post difference for never-treated state 2 is: 55-25=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 4 is 66-33= 33
* $DID_{4 vs 1,2} = 33-30=3$

**Contrast 3:**

* Pre-post difference for later-treated state 4 is: 39-22.5=16.5
* Pre-post difference for earlier-treated state 2 is: 36-16.5=19.5
* $DID_{3 vs 4} = 19.5-16.5=3$

**Contrast 4:**

* Pre-post difference for earlier-treated state 2 is: 60-36=24
* Pre-post difference for later-treated state 4 is: 66-39=27
* $DID_{4 vs 3} = 27-24=3$

```{r}
s5_mod <- lm(outcome ~  policy + state + factor(time), data = s5)
tidy(s5_mod)
```

The coefficient of the policy term is `r s5_mod$coefficients["policy1"]`
showing that the regression model estimates the ATT.

Using the `bacon()`function, we can see see the four comparisons being made and
the effect estimates for each comparison. The weight column in the output shows
how much weight each estimate contributes to the ATT. Here, the treated vs 
untreated comparisons are the most heavily weighted. 

```{r}
s5 %<>% mutate(state_n = as.numeric(as.character(state)),
               policy_n = as.numeric(as.character(policy)))
#bacon decomp says no problems
#only comparing treated vs untreated and the average effect estimate is 2
s5_bacon <- bacon(outcome ~ policy_n,
      data = s5,
      id_var = "state_n",
      time_var = "time")
```

We can also see a list of each comparison and it's weight: 

```{r}
s5_bacon
```

For pedagogical purposes, we also show the estimate if we applied the Callaway Sant’Anna estimator using the `att_gt()` function to estimate group-time ATTs.

```{r}
s5_cs <- att_gt(yname = "outcome", 
             tname = "time", 
             idname = "state_n", 
             gname = "time_first_treat", 
             data = s5, 
             anticipation = 0)

s5_cs_ag <- aggte(s5_cs, type="simple")
summary(s5_cs_ag)
```

Here, we see the ATT is equal to `r s5_cs_ag$overall.att` as estimated by the 
TWFE model.

### Scenario 6

Scenario 6 is similar to Scenario 5 except now the treatment effect increases
as time goes on, i.e., it is dynamic. The magnitude of the treatment effect is
the same for both treated groups.

```{r}
s6 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen6", 
                col_types = c("text", "numeric", "text", "numeric", "text", "numeric"))
```

```{r, echo=F}
ggplot(s6, aes(x = time, y = outcome)) + 
  geom_point(aes(col = factor(state))) + labs(title = "Scenario 6 (Dynamic)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2, col = "#005a32") + 
  geom_vline(aes(xintercept = 11.5), lty = 2, col = "#99000d") +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) +
  theme_bw()

```

```{r, echo=F}
s6_mean_1_4 <- s6 %>% filter(time < 5) %>%
  group_by(state) %>%
  summarise(mean_1_4 = mean(outcome))

s6_mean_5_20 <- s6 %>% filter(time >= 5) %>%
  group_by(state) %>%
  summarise(mean_5_20 = mean(outcome))

s6_DID1 <- bind_cols(s6_mean_1_4, s6_mean_5_20 %>% select(-state)) %>%
  mutate(time_diff = mean_5_20 - mean_1_4)


s6_mean_1_11 <- s6 %>% filter(time < 12) %>%
  group_by(state) %>%
  summarise(mean_1_11 = mean(outcome))

s6_mean_12_20 <- s6 %>% filter(time >= 12) %>%
  group_by(state) %>%
  summarise(mean_12_20 = mean(outcome))

s6_DID2 <- bind_cols(s6_mean_1_11, 
                     s6_mean_12_20 %>% select(-state)) %>%
  mutate(time_diff2 = mean_12_20 - mean_1_11)

s6_mean_5_11 <- s6 %>% filter(time >= 5 & time < 12) %>%
  group_by(state) %>%
  summarise(mean_5_11 = mean(outcome))

s6_DID3 <- bind_cols(s6_mean_1_4, 
                     s6_mean_5_11 %>% select(-state)) %>%
  mutate(time_diff3 = mean_5_11 - mean_1_4)

s6_DID4 <- bind_cols(s6_mean_5_11, 
                     s6_mean_12_20 %>% select(-state)) %>%
  mutate(time_diff4 = mean_12_20 - mean_5_11)
```

```{r, echo = F, fig.height = 8, fig.width = 10}
s6_comp1 <- ggplot(s6 %>% filter(state %in% c(1, 2, 3)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "A) Comparison 1 (Treated vs. never-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32")) +
  theme_bw() + 
  scale_x_continuous(limits = c(1, 20))  + 
  geom_label_repel(data = s6_DID1 %>% filter(state %in% c(1,2,3)), 
                  aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = s6_DID1 %>% filter(state %in% c(1,2,3)), 
                  aes(y = mean_5_20, x = 12.5, label = paste("mean_y:", mean_5_20), col = state))

s6_comp2 <- ggplot(s6 %>% filter(state %in% c(1, 2, 4)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "B) Comparison 2 (Treated vs. never-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s6_DID2 %>% filter(state %in% c(1,2,4)), 
                   aes(y = mean_1_11, x = 6, label = paste("mean_y:", mean_1_11), col = state)) +
  geom_label_repel(data = s6_DID2 %>% filter(state %in% c(1,2,4)), 
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s6_comp4 <- ggplot(s6 %>% filter(state %in% c(3, 4), time >= 5), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "C) Comparison 4 (Later vs. earlier-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s6_DID4 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state)) +
  geom_label_repel(data = s6_DID4 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s6_comp3 <- ggplot(s6 %>% filter(state %in% c(3, 4), time < 12), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "D) Comparison 3 (Earlier vs. later-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s6_DID3 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = s6_DID3 %>% filter(state %in% c(3,4)), 
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state))

s6_comp1 + s6_comp2 + s6_comp3 + s6_comp4 + plot_layout(guides = "collect")
```

**Contrast 1:**

* Pre-post difference for never-treated state 1 is: 43.5-13.5=30
* Pre-post difference for never-treated state 2 is: 44.5-14.5=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 3 is 57-16.5 = 40.5
* $DID_{3 vs 1,2} = 40.5-30 = 10.5$

**Contrast 2:**

* Pre-post difference for never-treated state 1 is: 54-24=30
* Pre-post difference for never-treated state 2 is: 55-25=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 4 is 70-33= 37
* $DID_{4 vs 1,2} = 37-30 = 7$

**Contrast 3:**

* Pre-post difference for later-treated state 4 is: 39-22.5=16.5
* Pre-post difference for earlier-treated state 3 is: 39-16.5=22.5
* $DID_{3 vs 4} = 22.5-16.5 = 6$

**Contrast 4:**

* Pre-post difference for earlier-treated state 3 is: 71-39=32
* Pre-post difference for later-treated state 4 is: 70-39=31
* $DID_{4 vs 3} = 31-32 = -1$

Contrasts 1 and 2 are okay because they are clean comparisons between never-treated
and treated states. Contrasts 3 and 4 are trickier. Contrast 3 compares the 
earlier treated state to the later treated state. It is okay because parallel
trends holds and it only uses the time before the later state is treated to 
inform the estimation. Contrast 4 is the problem -- it uses post-treatment time
from state 3 as the "pre" time for the comparison. The issue is that this post
time includes the effects of the treatment, and since the effect is dynamic it
does not satisfy the parallel trends assumption. However this contrast still 
contributes to the TWFE regression estimate!

Let's take a look...

```{r}
s6_mod <- lm(outcome ~ policy + state + factor(time), 
             data = s6)
tidy(s6_mod)
```
The coefficient of the policy term equals `r s6_mod$coefficients["policy1"] %>% round(2)`.
We can now use the Goodman Bacon decomposition to see how much weight each of the 
contrasts contributes to the effect estimate:

```{r}
s6 %<>% mutate(state_n = as.numeric(as.character(state)),
               policy_n = as.numeric(as.character(policy)))

s6_bacon <- bacon(outcome ~ policy_n,
      data = s6,
      id_var = "state_n",
      time_var = "time")
```
It's more helpful to view the four contrasts separately:

```{r}
s6_bacon
```
So we can see the estimates of the TE for each contrast are as we calculated 
them earlier. 

(WHAT DOES TE MEAN HERE?)

We can also confirm that you get the TWFE estimate by taking the weighted 
average of the Goodman-Bacon decomposition components:

```{r}
#see the TWFE estimate
sum(s6_bacon$estimate*s6_bacon$weight)
```
Okay, so we know the TWFE regression estimate is wrong, in particular because it is incorporating a 
contrast that we wouldn't want to make in practice between an earlier treated and a later treated group. To overcome this we can use the Callaway and Sant'Anna estimator. 

First, we running the `att_gt()` function and display its output using 
`summary()`. The ATT is estimated for each combination of treated state 
and time. It also includes estimates for pre-policy time, which helps with the 
evaluation of the parallel trends assumption or to see if there are any lead
effects ("anticipation") of the treatment on the outcome. 

```{r}
s6_cs <- att_gt(yname = "outcome", 
             tname = "time", 
             idname = "state_n", 
             gname = "time_first_treat", 
             data = s6, 
             anticipation = 0)

summary(s6_cs)
```

(I'M NOT SURE WHY THIS IS USEFUL TO SHOW)

Callaway and Sant'Anna provide many options for aggregating the group-time 
effects. The simplest option is specify `type = simple`. This estimate considered
only the clean contrasts of 1 and 2 and combines them by weighting by the total
time in the post-treatment period. For contrast 1, there are 16 
time periods post treatment and for contrast 2, there are 9 periods 
post-treatment. Thus this estimate is the weighted average: $10.5\times(16/25) + 7\times(9/25)=9.24$

```{r}
#aggregate the group time average treatment effects
#type = "simple": weighted average of all group-time average treatment effects
#with weights proportional to the group size
s6_cs_ag <- aggte(s6_cs, type="simple")
summary(s6_cs_ag)
```
Alternatively, we can specify `type = "group"` to get separate effect estimates 
according to time of implementation. Note that `group` denotes the time of 
the policy change for the treated states.

```{r}
s6_cs_ag2 <- aggte(s6_cs, type = "group")
summary(s6_cs_ag2)
```

Alternatively, you can estimate the dynamic effect separately for each time 
since the treatment has been introduced. Remember that this effect estimation
is also done for pre-treatment time, so don't be alarmed by all the rows in this 
table!

```{r}
s6_cs_ag3 <- aggte(s6_cs, type="dynamic")
summary(s6_cs_ag3)
```
In this case, it is very helpful to plot the effect estimates over time:

```{r}
ggdid(s6_cs_ag3)
```

For exposition's sake, we also show the estimate using `type = "calendar"`.
For this specific example, I don't love this aggregation -- I think for dynamic
effects it makes things a little wonky and doesn't reveal anything super helpful:

```{r}
#calendar time specific
#i don't love this one...but CS say they prefer it here
#https://bcallaway11.github.io/did/articles/did-basics.html
s6_cs_ag4 <- aggte(s6_cs, type = "calendar")
summary(s6_cs_ag4)
ggdid(s6_cs_ag4)
```

(I AGREE, NOT SURE WHAT IT ADDS AND IS CONFUSING. I'M GUESSING IT LOOKS THIS WAY BECAUSE OF THE TWO 
TREATED GROUPS HAVING DIFFERENT NUMBERS OF POST-TREATMENT OUTCOMES?)

### TWFE with `time_since_policy` specification

But what if we model TWFE with the `time_since_change` indicators?

```{r}
s6_mod2 <- lm(outcome ~ time_since_policy + state + factor(time), 
             data = s6)
tidy(s6_mod2) %>% filter(str_detect(term, "time_since")) %>% 
  mutate(time = as.numeric(gsub("time_since_policy", "", term))) %>% 
  arrange(time) 
```
In this setting, the regression model still works! The estimates from the policy
indicator variables equal the output from the Callaway Sant'Anna when specified 
using `type = "dynamic"`

**Takeway:** When the treatment effect is staggered and dynamic, you can still 
capture the effect estimate using a TWFE model so long as the policy effect is 
modeled using the `time since treatment` variable. 

(KEY QUESTION IS WHAT PARAMETER ARE YOU INTERESTED IN? DO YOU WANT A DYNAMIC ESTIMATE, 
OR ARE YOU MORE INTERESTED IN A SUMMARY PARAMETER? )

## Scenario 7

Scenario 7 is like Scenario 6 except now the treatment effect is dynamic AND
heterogeneous. In this scenario, the state that implements the policy change 
first has a stronger treatment effect (i.e., a steeper change in slope)

```{r}
s7 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen7", 
                col_types = c("text", "numeric", "text", "numeric", "text", "numeric"))
```

```{r, echo=F}
ggplot(s7, aes(x = time, y = outcome)) + 
  geom_point(aes(col = factor(state))) + labs(title = "Scenario 7 (Staggered, dynamic, and heterogeneous)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2, col = "#005a32") + 
  geom_vline(aes(xintercept = 11.5), lty = 2, col = "#99000d") +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) +
  theme_bw()
```
```{r, echo=F}
s7_mean_1_4 <- s7 %>% filter(time < 5) %>%
  group_by(state) %>%
  summarise(mean_1_4 = mean(outcome))

s7_mean_5_20 <- s7 %>% filter(time >= 5) %>%
  group_by(state) %>%
  summarise(mean_5_20 = mean(outcome))

s7_DID1 <- bind_cols(s7_mean_1_4, s7_mean_5_20 %>% select(-state)) %>%
  mutate(time_diff = mean_5_20 - mean_1_4)

s7_mean_1_11 <- s7 %>% filter(time < 12) %>%
  group_by(state) %>%
  summarise(mean_1_11 = mean(outcome))

s7_mean_12_20 <- s7 %>% filter(time >= 12) %>%
  group_by(state) %>%
  summarise(mean_12_20 = mean(outcome))

s7_DID2 <- bind_cols(s7_mean_1_11, 
                     s7_mean_12_20 %>% select(-state)) %>%
  mutate(time_diff2 = mean_12_20 - mean_1_11)

s7_mean_5_11 <- s7 %>% filter(time >= 5 & time < 12) %>%
  group_by(state) %>%
  summarise(mean_5_11 = mean(outcome))

s7_DID3 <- bind_cols(s7_mean_1_4, 
                     s7_mean_5_11 %>% select(-state)) %>%
  mutate(time_diff3 = mean_5_11 - mean_1_4)

s7_DID4 <- bind_cols(s7_mean_5_11, 
                     s7_mean_12_20 %>% select(-state)) %>%
  mutate(time_diff4 = mean_12_20 - mean_5_11)
```


```{r, echo = F, fig.height = 8, fig.width = 10}
s7_comp1 <- ggplot(s7 %>% filter(state %in% c(1, 2, 3)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "A) Contrast 1 (Treated vs. never-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32")) +
  theme_bw() + 
  scale_x_continuous(limits = c(1, 20))  + 
  geom_label_repel(data = s7_DID1 %>% filter(state %in% c(1,2,3)),
                  aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = s7_DID1 %>% filter(state %in% c(1,2,3)),
                  aes(y = mean_5_20, x = 12.5, label = paste("mean_y:", mean_5_20), col = state))

s7_comp2 <- ggplot(s7 %>% filter(state %in% c(1, 2, 4)), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "B) Contrast 2 (Treated vs. never-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s7_DID2 %>% filter(state %in% c(1,2,4)),
                   aes(y = mean_1_11, x = 6, label = paste("mean_y:", mean_1_11), col = state)) +
  geom_label_repel(data = s7_DID2 %>% filter(state %in% c(1,2,4)),
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s7_comp4 <- ggplot(s7 %>% filter(state %in% c(3, 4), time >= 5), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "C) Contrast 4 (Later vs. earlier-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 11.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw()+ 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s7_DID4 %>% filter(state %in% c(3,4)),
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state)) +
  geom_label_repel(data = s7_DID4 %>% filter(state %in% c(3,4)),
                   aes(y = mean_12_20, x = 16, label = paste("mean_y:", mean_12_20), col = state))

s7_comp3 <- ggplot(s7 %>% filter(state %in% c(3, 4), time < 12), aes(x = time, y = outcome)) +   
  geom_point(aes(col = factor(state))) + labs(title = "D) Contrast 3 (Earlier vs. later-treated)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2) +
  scale_color_manual(values = c("#005a32", "#99000d")) +
  theme_bw() + 
  scale_x_continuous(limits = c(1, 20)) +
  geom_label_repel(data = s7_DID3 %>% filter(state %in% c(3,4)),
                   aes(y = mean_1_4, x = 2.5, label = paste("mean_y:", mean_1_4), col = state)) +
  geom_label_repel(data = s7_DID3 %>% filter(state %in% c(3,4)),
                   aes(y = mean_5_11, x = 8, label = paste("mean_y:", mean_5_11), col = state))

s7_comp1 + s7_comp2 + s7_comp3 + s7_comp4 + plot_layout(guides = "collect")
```

**Contrast 1:**

* Pre-post difference for never-treated state 1 is: 43.5-13.5=30
* Pre-post difference for never-treated state 2 is: 44.5-14.5=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 3 is 64.5-16.5 = 48
* $DID_{3 vs 1,2} = 48-30 = 18$

**Contrast 2:**

* Pre-post difference for never-treated state 1 is: 54-24=30
* Pre-post difference for never-treated state 2 is: 55-25=30
* Implying the average pre-post difference equals 30 for the never-treated states
* Pre-post difference for treated state 4 is 68-33= 35
* $DID_{4 vs 1,2} = 35-30 = 5$

**Contrast 3:**

* Pre-post difference for later-treated state 4 is: 39-22.5=16.5
* Pre-post difference for earlier-treated state 3 is: 42-16.5=25.5
* $DID_{3 vs 4} = 25.5-16.5 = 9$

**Contrast 4:**

* Pre-post difference for earlier-treated state 3 is: 82-42=40
* Pre-post difference for later-treated state 4 is: 68-39=29
* $DID_{4 vs 3} = 29-40 = -11$

Similar to the previous scenario, parallel trends is satisfied for contrasts 1-3, 
so we can be okay with these contrasts contributed to the average effect estimate. 
Like last time, contrast 4 is the problem, and here the problem is intensified
because the "control" state -- which was treated in an earlier period -- is 
undergoing a large treatment effect in the "pre" period (times 5-11). This 
makes the pre-post different in the control state larger than the pre-post difference
in the comparison state -- because the control state is still experiencing a stronger
(i.e., steeper slope) treatment effect than the state that is treated at time 12. 
This makes the DID for this group and time look negative -- like the treatment
is associated with a decrease in the outcome!

Let's take a look at the regression output:

```{r}
s7_mod <- lm(outcome ~ policy + state + factor(time), 
             data = s7)
tidy(s7_mod)
```
The coefficient of the policy term equals `r s7_mod$coefficients["policy1"] %>% round(2)`.

(CAN WE SAY WHAT THE COEFFICIENT SHOULD BE?)

We can now use the Goodman Bacon decomposition to see how much weight each of 
the contrasts contributes to the effect estimate:

```{r}
s7 %<>% mutate(state_n = as.numeric(as.character(state)),
               policy_n = as.numeric(as.character(policy)))

s7_bacon <- bacon(outcome ~ policy_n,
      data = s7,
      id_var = "state_n",
      time_var = "time")
```

```{r}
s7_bacon
```

```{r}
#see the TWFE estimate
sum(s7_bacon$estimate*s7_bacon$weight)
```
Now that we know how to use the Callaway and Sant'Anna code, we can apply it to
this example as well. Let's start by running the `att_gt()` function, remembering
that the output is for every combination of group and time:

```{r}
s7_cs <- att_gt(yname = "outcome", 
             tname = "time", 
             idname = "state_n", 
             gname = "time_first_treat", 
             data = s7, 
             anticipation = 0)

summary(s7_cs)
```

Let's aggregate these ATTs using `type = simple`. This estimate considered
only the clean contrasts of 1 and 2 and combines them by weighting by the total
time in the post-treatment period. For contrast 1, there are 16 
time periods post treatment and for contrast 2, there are 9 periods 
post-treatment. Thus this estimate is the weighted average: $18\times(16/25) + 5\times(9/25)=13.32$

```{r}
#aggregate the group time average treatment effects
#type = "simple": weighted average of all group-time average treatment effects
#with weights proportional to the group size
s7_cs_ag <- aggte(s7_cs, type="simple")
summary(s7_cs_ag)
```
Alternatively, we can specify `type = "group"` to get separate effect estimates 
according to time of implementation and see that these match our by-hand calculations: 

```{r}
s7_cs_ag2 <- aggte(s7_cs, type = "group")
summary(s7_cs_ag2)
```

Alternatively, you can estimate the dynamic effect separately for each time 
since the treatment has been introduced. Remember that this effect estimation
is also done for pre-treatment time, so don't be alarmed by all the rows in this 
table!

```{r}
s7_cs_ag3 <- aggte(s7_cs, type="dynamic")
summary(s7_cs_ag3)
```
In this case, it is very helpful to plot the effect estimates over time:

```{r}
ggdid(s7_cs_ag3)
```

For exposition's sake, let's also show the estimate using `type = "calendar"`.
For this specific example, I don't love this aggregation -- I think for dynamic
effects it makes things a little wonky and doesn't reveal anything super helpful:

```{r}
#calendar time specific
#i don't love this one...but CS say they prefer it here
#https://bcallaway11.github.io/did/articles/did-basics.html
s7_cs_ag4 <- aggte(s7_cs, type = "calendar")
summary(s7_cs_ag4)
ggdid(s7_cs_ag4)
```


### TWFE with `time_since_policy` specification

But what if model TWFE with the `time_since_change` indicators? Here I run the 
regression and pull out the effect estimates since the regression output is so large

```{r}
s7_mod2 <- lm(outcome ~ time_since_policy + state + factor(time), 
             data = s7)
tidy(s7_mod2) %>% filter(str_detect(term, "time_since")) %>% 
  mutate(time = as.numeric(gsub("time_since_policy", "", term))) %>% 
  arrange(time) 
```
The effect estimates on the policy indicators do not equal an average of the ATTs
from the two states for each time period. For this example, this specification 
gives estimates that are systematically lower than those produced by the Callaway
Sant'Anna model. 

**Takeway:** When the treatment effect is staggered, dynamic, and hetergeneous, 
the TWFE regression specification will produced biased results, as will the 
specification using a categorical variable for time since treatment. In this 
case, it is best to use another approach like the Callaway Sant'Anna function.

## To Do

* Explain how the weight is calculated in the Bacon decomp -- when would it place 
a lot of weight on the forbidden contrasts? I have another R file called Bacon
weights where I try and implement the weight formulas but I get stuck...
* When will the weight be negative? I still haven't figured out what scenario this
corresponds to. Possibly in a setting where there are no ever-treated states?
* Need to say something about standard error estimation - the code as written is not concerned
with SE at all. So may want to say something or modify the code.  
* CONSIDER IMPLEMENTING THE STAGGERED REGRESSION APPRAOCH
* DISCUSSION OF DIFFERENT TARGET PARAMETERS, AND WHY SOME MIGHT BE OF MORE INTEREST THAN OTHERS
* https://causalinf.substack.com/p/illustrating-the-bias-of-twfe-in?utm_source=url
maybe try implementing this example

```{r}
# s8 <- read_xlsx(here("data", "scenarios.xlsx"), sheet = "scen8", 
#                 col_types = c("text", "text", "text", "numeric"))
# 
# s8 %<>% mutate(state_n = as.numeric(as.character(state)),
#                policy_n = as.numeric(as.character(policy)))

# s8_bacon <- bacon(formula = outcome ~ policy_n, 
#                   data = s8,
#                   id_var = "state_n",
#                   time_var = "time")
```
```

```{r, echo=F}
ggplot(s7, aes(x = time, y = outcome)) + 
  geom_point(aes(col = factor(state))) + labs(title = "Scenario 7 (Staggered, dynamic, and heterogeneous)") +
  geom_line(aes(col = factor(state))) + 
  geom_vline(aes(xintercept = 4.5), lty = 2, col = "#005a32") + 
  geom_vline(aes(xintercept = 11.5), lty = 2, col = "#99000d") +
  scale_color_manual(values = c("#a6cee3","#1f78b4","#005a32", "#99000d")) +
  theme_bw()
```

